---
title: "DDS Case Study 2 Analysis"
author: "Akbar Thobani"
date: "04/04/2020"
output:
  pdf_document: default
  html_document: default
keep_md: yes
---

#Problem Introduction
DDS Analytics has tasked us with to uncover insights and trends specific to Job Roles within the company and how those factors contribute to turnover/attrition. 


#Analysis Objectives
1) Uncover any intersting trends specific to Job Role <br />
2) Report the top 3 factors that contribute to turnover <br />
3) Build a model to predict attrition<br />
4) Build a model to predict salary <br />

#Packages
```{r message=FALSE}
library(ggplot2)
library(corrplot)
library(dplyr)
library(caret)
library(MASS)
library(randomForest)
library(e1071)
library(tidyverse)
```

#Data Import
```{r echo=TRUE}
dfTrain <- read.csv(file="CaseStudy2-data.csv", header=TRUE, stringsAsFactors=TRUE)
dfVal <- read.csv(file="CaseStudy2validation.csv", header=TRUE, stringsAsFactors=TRUE)

dfCompAtt <- read.csv(file="CaseStudy2CompSet No Attrition.csv", header=TRUE, stringsAsFactors=TRUE)
dfCompSal <- read.csv(file="CaseStudy2CompSet No Salary.csv", header=TRUE, stringsAsFactors=TRUE)
```

#Data Check
We want to make sure the dataset provided does not have any missing values or mixed data types before we begin our explaoratory and modeling exercises. 
```{r echo=TRUE}
str(dfVal)

#No missing values 
#sum(is.na(dfVal))
#colSums(is.na(dfTrain))
#colSums(is.na(dfVal))

#View(summary(df))
```


#Data Preparation
There are a few variables that seem useless for the purposes of this analysis. ID, Standard Hours, Employee Number and Employee Count will be removed from the table
```{r echo=TRUE}
#Recode Attrition Column to numeric if necessary
#df$Attrition2 <- ifelse(df$Attrition == "Yes", 1, 0)

#Drop ID, StandardHours, EmployeeCount, Over18 columns
#Most values do not change so SD is 0

df_stage <- dfTrain[,!(names(dfTrain) %in% c("ID", "StandardHours", "EmployeeNumber", "EmployeeCount", "Over18"))]
df_val_stage <- dfVal[,!(names(dfVal) %in% c("ID", "StandardHours", "EmployeeNumber", "EmployeeCount", "Over18"))]

```

An additional dataframe was created with only numeric values to be read by a correlation heatmap later in the analysis.
```{r echo=TRUE}
#Return numeric values only
df_numeric <- df_stage[, sapply(df_stage, is.numeric)]
df_val_numeric <- df_val_stage[, sapply(df_val_stage, is.numeric)]

#Correlation Plot
df_corr <- round(cor(df_numeric),2)
```


#Data Exploration
More than 80% of the training dataset consist of employees who are still retained
```{r echo=TRUE}
table(dfTrain$Attrition)
barplot(prop.table(table(dfTrain$Attrition)))
```

We created a correlation matrix to view possible multicollinearity between variables that need to be addressed before the modeling phase to avoid redundancy. 

The variables below seem to have high collinearity so we will remove some of them for the Custom model at a later phase:

MonthlyIncome corr JobLevel <br />
PercentSalaryHike corr PerformanceRating <br />
TotalWorkingYears corr JobLevel <br />
Age corr TotalWorkingYears <br />
YearsInCurrentRole corr TotalWorkingYears 
```{r echo=TRUE}
corrplot(df_corr, order="FPC", title="Variable Corr Heatmap",tl.srt=45,method = "pie")
```

Figure 1 shows the count distribution of employees by job role. Sales Executive jobs are the most prevalent at 22% of all Job Roles followed closely by Research Scientist at 20% and Lab Technicians at 18% rounding out the top 3.
```{r echo=TRUE}
p1 <- ggplot(dfTrain, aes(x=JobRole), color=JobRole) + ggtitle("Figure 1: Job Role") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..), fill=JobRole),  width = 0.7) + 
  labs(y="Percentage") + 
  coord_flip() + 
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.7)) 
p1
```

Figure 2 is another view showing the percentage of total by each Job Role
```{r echo=TRUE}
p2 <- ggplot(dfTrain, aes(x=JobRole, y = ..prop.., group=1)) +
  geom_bar() +
  geom_text(stat = "count", 
            aes(label = round(..prop.., 2), y = ..prop.. + 0.02)) +
            coord_flip() +
  ggtitle("Figure 2: Job Role by Percent of Total") +
  theme(plot.title = element_text(hjust = 0.5)) 
p2
```

Although Lab Technician roles account for almost 17% of all Job Roles, the attrition total and porportion seems to abnormally high. 
```{r echo=TRUE}
p3 <- ggplot(dfTrain,aes(x = JobRole,fill = Attrition)) +
    geom_bar(position = position_stack(reverse = FALSE)) +
    ggtitle("Job Role vs Attrition - Count") +
    coord_flip() +
    theme(legend.position = "top") +
    theme(plot.title = element_text(hjust = 0.8))

p3
```

Naive Bayes Model
```{r echo=TRUE}
Naive_Bayes_Model=naiveBayes( Attrition~., data=dfTrain)

#Summary of model
Naive_Bayes_Model
#Prediction on validation dataset
dfPreds0=predict(Naive_Bayes_Model,dfVal)
confusionMatrix(table(dfPreds0,dfVal$Attrition))
cMatrix<-table(dfPreds0, dfVal$Attrition)
plot(cMatrix, col="blue", ylab="Actual", xlab="Predicted", main='Naive Bayes Confusion Matrix')

#Test against competition dataset
dfPreds_Comp_Att=predict(Naive_Bayes_Model,dfCompAtt)
```

KNN nearest neighbor classification
```{r echo=TRUE}

indx <- sapply(dfTrain, is.factor)
dfTrain[indx] <- lapply(dfTrain[indx], function(x) as.numeric(as.factor(x)))
dfTrain <- dfTrain[, sapply(dfTrain, is.numeric)]

indx <- sapply(dfVal, is.factor)
dfVal[indx] <- lapply(dfVal[indx], function(x) as.numeric(as.factor(x)))
dfVal <- dfVal[, sapply(dfVal, is.numeric)]

indx <- sapply(dfCompAtt, is.factor)
dfCompAtt[indx] <- lapply(dfCompAtt[indx], function(x) as.numeric(as.factor(x)))
dfCompAtt <- dfCompAtt[, sapply(dfCompAtt, is.numeric)]

# k = 10
#classifications = knn(dfTrain[,c(4:35)],dfVal[,c(4:35)],dfTrain$Attrition,k = 9, l=0, prob = FALSE, use.all = TRUE)
#table(dfVal$Attrition,classifications)
#confusionMatrix(table(dfVal$Attrition,classifications))

```


#Output Dataset
```{r echo=FALSE}
#Write out Competition results
dfPreds1=cbind(dfCompAtt$ID, dfPreds_Comp_Att)
dfPreds=as.data.frame(dfPreds1)
names(dfPreds)<- c("ID", "Attrition")
write.csv(dfPreds,"DDSCaseStudy2_Predict_Attrition.csv",row.names = FALSE)

```

Linear Regression Model to Predict Salary
```{r echo=TRUE}
dfTrain <- read.csv(file="CaseStudy2-data.csv", header=TRUE, stringsAsFactors=TRUE)

lr_mod <- lm(MonthlyIncome ~ Attrition + JobSatisfaction + Age + Education + JobLevel +
  PerformanceRating + YearsSinceLastPromotion + YearsWithCurrManager + WorkLifeBalance +
  YearsAtCompany + JobRole + EducationField , data=dfTrain)

predictMonthlyIncome <- predict(lr_mod, dfCompSal, se.fit = TRUE)
summary(lr_mod)
plot(lr_mod)

```  

#Output Monthly Income Dataset
```{r echo=FALSE}
#Write out Competition results
dfIncomePreds=cbind(dfCompSal$ID, round(predictMonthlyIncome$fit,0))
dfIncomePreds=as.data.frame(dfIncomePreds)
names(dfIncomePreds)<- c("ID", "Monthly Income")
write.csv(dfIncomePreds,"DDSCaseStudy2_Predict_MonthlyIncome.csv",row.names = FALSE)

```
